Nail: A practical interface generator for data formats

### 2014-11-25 paper/login/eval.tex
pass

First, we ask whether Nail can handle real-world data formats and whether its techniques are
critical to that. We used Nail to implement grammars
for seven protocols with a range of challenging features.
Figure~\ref{fig:eval-protocols} summarizes our results.
We find that despite the challenging aspects of these protocols, Nail is
able to capture the protocols, by relying on its novel features: dependent
fields, streams, and transforms.  In contrast, state-of-the-art parser
generators would be unable to fully handle 5 out of the 7 data formats.
In the rest of this subsection, we describe the DNS and Zip grammars in
more detail, focusing on how Nail's features enable us to support these
formats.

@@

\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format, as specified
by PKWare.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an
\emph{end-of-directory header}. This header contains a variable-length comment, so it has to be
located by scanning backwards from the end of the file until a magic number and a valid length field
is found. Many ZIP implementations disagree on the exact semantics of this, such as when the comment
contains the magic number~\cite{wolf:berlinsides-zip}.

This header  contains points to the \emph{ZIP directory}, contains an entry every file
in the archive.  Each entry stores file metadata in addition to the
offset of a \emph{local file header}. The local file header duplicates
most information from the directory entry header and is followed immediately by the compressed
archive entry.
Duplicating information made sense when ZIP files were stored on floppy
disks with slow seek times and high fault rates, and memory constraints
made it impossible to keep the ZIP directory in memory.

 However, if attackers craft inconsistent files, hand written parsers fail in interesting ways, e.g. 
the length in the central directory is used to allocate memory and the
length in the local directory is used to extract without checking that
they are equal first, as occured in Python bug 20078.

Nail avoids these issues by capturing these redundancies with dependent fields and decompressing
files transparently with transformations, and the programmer
is presented with a data structure containing the archive contents. To handle formats based on ZIP,
such as office documents, our ZIP grammar could be extended to apply additional grammars to the
files within an archive.

@@

Our DNS server parses a zone file, listens to incoming DNS requests,
parses them, and generates appropriate responses.  The DNS server is
implemented in 183 lines of C, together with 48 lines of Nail grammar
and 64 lines of C code implementing stream transforms for DNS label
compression.  In comparison, Hammer~\cite{hammer-parser} ships with a toy
DNS server that responds to any valid DNS query with a CNAME record to the
domain ``spargelze.it''.  Their server consists of 683 lines of C, mostly
custom validators, semantic actions, and data structure definitions, with
 52 lines of code defining the grammar with Hammer's combinators.
Their DNS server does not implement label compression, zone files, etc.
From this, we conclude that Nail leads to much more compact code for
dealing with DNS packet formats.

We ran the DNS fuzzer provided with the Metasploit
framework on our DNS server, which sent randomly
corrupted DNS queries to our server for 4 hours, during which it did
not crash or trigger the stack or heap corruption detector.

@@

In our full paper~\cite{bangert:nail-osdi14}, we present a study of 15 ZIP parsing bugs.
11 of these vulnerabilities involved memory corruption during input handling, which generated code
is immune to by design.
More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.
Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
that caused a vulnerability in the Python ZIP library. The other three vulnerabilities exist
because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
parsers are not very reusable, as they build application-specific data structures and are tightly
coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
avoiding protocol misunderstandings.

>>>>

First, we'll demonstrate how Nail's features help process real-world protocols. We used Nail to implement grammars
for seven protocols with a range of challenging features.
Figure~\ref{fig:eval-protocols} summarizes our results.
We find that despite the challenging aspects of these protocols, Nail is
able to capture the protocols, by relying on its novel features: dependent
fields, streams, and transforms.  In contrast, state-of-the-art parser
generators would be unable to fully handle 5 out of the 7 data formats.

@@

\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format, as specified
by PKWare. At the end of each ZIP file is an
\emph{end-of-directory header}. This header contains a variable-length comment, so it has to be
located by scanning backwards from the end of the file until a magic number and a valid length field
is found. Many ZIP implementations disagree on the exact semantics of this, such as when the comment
contains the magic number~\cite{wolf:berlinsides-zip}.
This header  contains points to the \emph{ZIP directory}, contains an entry every file
in the archive.  Each entry stores file metadata in addition to the
offset of a \emph{local file header}. The local file header duplicates
most information from the directory entry header and is followed immediately by the compressed
archive entry.
Duplicating information made sense when ZIP files were stored on floppy
disks with slow seek times and high fault rates, but nowadays lead to parsers being confused, such
as in the recent Android Master Key bug.

%  However, if attackers craft inconsistent files, hand written parsers fail in interesting ways, e.g. 
% the length in the central directory is used to allocate memory and the
% length in the local directory is used to extract without checking that
% they are equal first, as occured in Python bug 20078.

Nail avoids these issues by capturing these redundancies with dependent fields and decompressing
files transparently with transformations, and the programmer
is presented with a data structure containing the archive contents. Nail even allows parsing the
contents of an archive file -- allowing formats based on ZIP, such as Microsoft Office documents, to
be handled with one grammar.

@@

Our DNS server parses a zone file, listens to incoming DNS requests,
parses them, and generates appropriate responses.  The DNS server is
implemented in 183 lines of C, together with 48 lines of Nail grammar
and 64 lines of C code implementing stream transforms for DNS label
compression.  In comparison, Hammer~\cite{hammer-parser} ships with a toy
DNS server that responds to any valid DNS query with a CNAME record to the
domain ``spargelze.it''.  Their server consists of 683 lines of C, mostly
custom validators, semantic actions, and data structure definitions, with
 52 lines of code defining the grammar with Hammer's combinators.
Their DNS server does not implement label compression, zone files, etc.

@@

In our full paper~\cite{bangert:nail-osdi14}, we present a study of 15 ZIP parsing bugs.
11 of these vulnerabilities involved memory corruption during input handling, which generated code
is immune to by design. We also fuzz-tested our DNS server.
More interestingly, Nail also protects against parsing inconsistency vulnerabilities like the four others we studied.
Nail grammars explicitly encode duplicated information such as the redundant length fields in ZIP
that caused a vulnerability in the Python ZIP library. The other three vulnerabilities exist
because multiple implementations of the same protocol disagree on some inputs. Hand-written protocol
parsers are not very reusable, as they build application-specific data structures and are tightly
coupled to the rest of the code. Nail grammars, however, can be re-used between applications,
avoiding protocol misunderstandings.
